Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-09-09 submission/manuscript.Rmd
fig 4 legend

\begin{figure}
  \centering 
  \includegraphics{../submission/compiled_figs/EpiLCs_too.pdf}
  \caption[\textit{Kdm5c}-KO epiblast-like cells express key drivers of germline identity ]{\textbf{\textit{Kdm5c}-KO epiblast-like cells express key drivers of germline identity A.} Top - Diagram of \textit{in vivo} differentiation of embryonic stem cells (ESCs) of the inner cell mass into epiblast stem cells. Bottom - \textit{in vitro} differentiation of ESCs into epiblast-like cells (EpiLCs). \textbf{B.} Representative images of wild-type (WT) and \textit{Kdm5c}-KO cells during ESC to EpiLC differentiation. Brightfield images taken at 20X. \textbf{C.} No significant difference in primed pluripotency maker expression in wild-type versus \textit{Kdm5c}-KO EpiLCs. Welch's t-test, expression in transcripts per million (TPM). \textbf{D.} Number of tissue-enriched differentially expressed genes (DEGs) in \textit{Kdm5c}-KO EpiLCs. * p<0.05, ** p<0.01, *** p<0.001, Fisher's exact test. \textbf{E.} Upset plot displaying the overlap of germline DEGs expressed in \textit{Kdm5c}-KO EpiLCs, amygdala (AMY), and hippocampus (HIP) RNA-seq datasets. \textbf{F.} Average bigwigs of an example germline gene, \textit{Cyct}, that is dysregulated \textit{Kdm5c}-KO EpiLCs (top, purple), amygdala (middle, red), and hippocampus (bottom, blue). \textbf{G.} enrichPlot comparing enriched biological process gene ontologies for \textit{Kdm5c}-KO EpiLC, amygdala, and hippocampus germline DEGs. \textbf{H.} Top left - Example germline identity DEGs unique to EpiLCs. Top right - Example 2-cell genes that are not dysregulated in \textit{Kdm5c}-KO EpiLCs. p-values for Welch's t-test. Bottom - Average bigwigs of \textit{Dazl} and \textit{Stra8} expression in wild-type and \textit{Kdm5c}-KO EpiLCs.}
  \label{figurelabel}
\end{figure}

\begin{figure}
  \centering 
  \includegraphics{../submission/compiled_figs/EpiLC_XXvsXY.pdf}
  \caption[Chromosomal sex influences \textit{Kdm5c}-KO germline gene misexpression.]{\textbf{Chromosomal sex influences \textit{Kdm5c}-KO germline gene misexpression.} \textbf{A-B.} Expression of tissue-enriched genes (Li et al 2017) in the male \textit{Kdm5c}-KO amygdala (A) and hippocampus (B). Left - MA plot of mRNA-sequencing. Right - Number of tissue-enriched differentially expressed genes (DEGs). * p<0.05, ** p<0.01, *** p<0.001, Fisher's exact test \textbf{C.} Left - Average bigwigs of an example aberrantly expressed testis-enriched DEG, \textit{FK506 binding protein 6 (Fkbp6)} in the wild-type (WT) and \textit{Kdm5c}-KO (5CKO) amygdala (red) and hippocampus (teal). Right - Expression of \textit{Cyct} in wild-type tissues from NCBI Gene, with testis highlighted in blue and brain tissues highlighted in red. \textbf{D.} Left - Average bigwigs of an example ovary-enriched DEG, \textit{Zygotic arrest 1 (Zar1)}. Right - Expression of \textit{Zar1} in wild-type tissues from NCBI Gene, with ovary highlighted in teal and brain tissues highlighted in red. \textbf{E.} Left - Average bigwigs of an example liver-enriched DEG, \textit{Apolipoprotein C-I (Apoc1)}. Right - Expression of \textit{Apoc1} in wild-type tissues from NCBI Gene, with liver highlighted in orange and brain tissues highlighted in red.}
  \label{figurelabel}
\end{figure}

>>>>

\begin{figure}
  \centering 
  \includegraphics{../submission/compiled_figs/EpiLCs_too.pdf}
  \caption[\textit{Kdm5c}-KO epiblast-like cells express key drivers of germline identity ]{\textbf{\textit{Kdm5c}-KO epiblast-like cells express key drivers of germline identity A.} Top - Diagram of \textit{in vivo} differentiation of embryonic stem cells (ESCs) of the inner cell mass into epiblast stem cells. Bottom - \textit{in vitro} differentiation of ESCs into epiblast-like cells (EpiLCs). \textbf{B.} Representative images of wild-type (WT) and \textit{Kdm5c}-KO cells during ESC to EpiLC differentiation. Brightfield images taken at 20X. \textbf{C.} No significant difference in primed pluripotency maker expression in wild-type versus \textit{Kdm5c}-KO EpiLCs. Welch's t-test, expression in transcripts per million (TPM). \textbf{D.} Number of tissue-enriched differentially expressed genes (DEGs) in \textit{Kdm5c}-KO EpiLCs. * p<0.05, ** p<0.01, *** p<0.001, Fisher's exact test. \textbf{E.} Upset plot displaying the overlap of germline DEGs expressed in \textit{Kdm5c}-KO EpiLCs, amygdala (AMY), and hippocampus (HIP) RNA-seq datasets. \textbf{F.} Average bigwigs of an example germline gene, \textit{Cyct}, that is dysregulated \textit{Kdm5c}-KO EpiLCs (top, purple), amygdala (middle, red), and hippocampus (bottom, blue). \textbf{G.} enrichPlot gene ontology analysis comparing enriched biological processes for \textit{Kdm5c}-KO EpiLC, amygdala, and hippocampus germline DEGs. \textbf{H.} Top left - Example germline identity DEGs unique to EpiLCs. Top right - Example 2-cell genes that are not dysregulated in \textit{Kdm5c}-KO EpiLCs. p-values for Welch's t-test. Bottom - Average bigwigs of \textit{Dazl} and \textit{Stra8} expression in wild-type and \textit{Kdm5c}-KO EpiLCs.}
  \label{figurelabel}
\end{figure}

\begin{figure}
  \centering 
  \includegraphics{../submission/compiled_figs/EpiLC_XXvsXY.pdf}
  \caption[Chromosomal sex influences \textit{Kdm5c}-KO germline gene misexpression.]{\textbf{Chromosomal sex influences \textit{Kdm5c}-KO germline gene misexpression.} \textbf{A.} Total number of germline-enriched RNA-seq DEGs for male hemizygous \textit{Kdm5c} knockout EpiLCs (XY 5CKO, purple), female heterozygous \textit{Kdm5c} knockout (XX 5CHET, orange), female homozygous \textit{Kdm5c} knockout (XX 5CKO, light orange) EpiLCs. \textbf{B.} Left - Eulerr overlap of \textit{Kdm5c} mutant male and female EpiLC germline DEGs. Right - Example of germline DEGs unique to females or males, \textit{Tktl2} and \textit{Esx1}.\textbf{C.} enrichPlot gene ontology analysis comparing enriched biological processes for germline DEGs shared between \textit{Kdm5c} mutant males and females (Shared), or unique to one sex (XX only or XY only). \textbf{D.} Heatmap of germline DEGs shared between male and female mutants. Color is the average log 2 fold change from sex-matched wild-type. \textbf{E.} Number of genes within each cluster from D. Clusters with higher expression in females compared to males (XX-biased) highlighted in orange. \textbf{F.} Example average bigwigs of a male and female shared germline DEG \textit{D1Pas1} that is more highly expressed in female mutants. \textbf{G.} Left - Number of all female germline DEGs located on each chromosome over the total number of germline-enriched genes on that chromsome. Right - Percentage of germline DEGs that lie on each chromosome for each \textit{Kdm5c} mutant. X chromosome highlighted in black. \textbf{H.} Sankey diagram classifying egg-biased (pink) and sperm-biased (blue) and unbiased (purple) mouse germline-enriched genes. \textbf{I.} Number of egg, sperm, or unbiased germline DEGs for male and female \textit{Kdm5c} mutants. \textbf{J.} Example bigwigs of egg-biased (\textit{Zp3}), sperm-biased (\textit{Dkkl1}), and unbiased (\textit{Dazl}) dysregulated in both male and female \textit{Kdm5c} mutants.}
  \label{figurelabel}
\end{figure}
